---
title: "LAB 10"
author: "Caroline He"
date: "11/8/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Packages
```{r}
library(RSQLite)
library(DBI)
library(sqldf)
```

## Setup
```{r}
# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
actor <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/actor.csv")
rental <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/rental.csv")
customer <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/customer.csv")
payment <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/payment_p2007_01.csv")

# Copy data.frames to database
dbWriteTable(con, "actor", actor)
dbWriteTable(con, "rental", rental)
dbWriteTable(con, "customer", customer)
dbWriteTable(con, "payment", payment)
```

```{r}
dbListTables(con)
```

```{sql, connection=con, output.var="x1"}
PRAGMA table_info(actor)
```

This is equivalent to use `dbGetQuery`
```{r}
dbGetQuery(con, "PRAGMA table_info(actor)")
```

# Exercise 1
#### Retrive the actor ID, first name and last name for all actors using the actor table. Sort by last name and then by first name.
```{r}
dbGetQuery(con, "SELECT actor_id, first_name, last_name
FROM actor /* you can add commit using
mutiple lines! */
ORDER by last_name, first_name 
LIMIT 5")
```

# Exercise 2
#### Retrive the actor ID, first name, and last name for actors whose last name equals ‘WILLIAMS’ or ‘DAVIS’.
```{r}
dbGetQuery(con, "SELECT actor_id, first_name, last_name
FROM actor 
WHERE last_name IN ('WILLIAMS', 'DAVIS')")
```

# Exercise 3
#### Write a query against the rental table that returns the IDs of the customers who rented a film on July 5, 2005 (use the rental.rental_date column, and you can use the date() function to ignore the time component). Include a single row for each distinct customer ID.
```{r}
dbGetQuery(con, "PRAGMA table_info(rental)")
```

```{r}
dbGetQuery(con," 
SELECT DISTINCT customer_id 
FROM rental
WHERE date(rental_date) = '2005-07-05' LIMIT 5")
```

# Exercise 4
### Exercise 4.1
#### Construct a query that retrives all rows from the payment table where the amount is either 1.99, 7.99, 9.99.
```{r}
q <- dbSendQuery(con, "SELECT *
FROM payment
WHERE amount IN (1.99, 7.99, 9.99)")
dbFetch(q, n = 10)
dbFetch(q, n = 10)
dbClearResult(q)
```

### Exercise 4.2
#### Construct a query that retrives all rows from the payment table where the amount is greater then 5
```{r}
dbGetQuery(con, "SELECT *
FROM payment
WHERE amount > 5 LIMIT 5")
```

# Exercise 5
#### Retrive all the payment IDs and their amount from the customers whose last name is ‘DAVIS’.
```{r}
dbGetQuery(con, "SELECT p.payment_id, p.amount
FROM payment AS p
INNER JOIN customer AS c ON p.customer_id=c.customer_id
WHERE c.last_name = 'DAVIS'")
```

# Exercise 6
### Exercise 6.1
#### Use COUNT(*) to count the number of rows in rental
```{r}
dbGetQuery(con, "
SELECT customer_id, 
COUNT(*) AS 'Rentals'
FROM payment
")
```

### Exercise 6.2
#### Use COUNT(*) and GROUP BY to count the number of rentals for each customer_id
```{r}
dbGetQuery(con, "
SELECT customer_id, COUNT(*) AS 'Rentals'
FROM rental GROUP BY customer_id LIMIT 10
")
```

### Exercise 6.3
#### Repeat the previous query and sort by the count in descending order
```{r}
dbGetQuery(con, "
SELECT customer_id, COUNT(*) AS 'Rentals'
FROM rental GROUP BY customer_id
ORDER BY `Rentals` DESC LIMIT 10
")
```

### Exercise 6.4
#### Repeat the previous query but use HAVING to only keep the groups with 40 or more.
```{r}
dbGetQuery(con, "
SELECT customer_id, COUNT(*) AS 'Rentals'
FROM rental GROUP BY customer_id
HAVING `Rentals` >= 40
ORDER BY `Rentals` DESC
")
```


# Exercise 7
#### The following query calculates a number of summary statistics for the payment table using MAX, MIN, AVG and SUM
```{r}
dbGetQuery(con, "
SELECT 
  MAX(amount) AS `max`,
  MIN(amount) AS `min`,
  AVG(amount) AS `avg`,
  SUM(amount) AS `sum`
FROM payment")
```

### Exercise 7.1
#### Modify the above query to do those calculations for each customer_id
```{r}
dbGetQuery(con, "
SELECT 
  customer_id,
  MAX(amount) AS `max`,
  MIN(amount) AS `min`,
  AVG(amount) AS `avg`,
  SUM(amount) AS `sum`
FROM payment GROUP BY customer_id
LIMIT 5")
```

### Exercise 7.2
#### Modify the above query to only keep the customer_ids that have more then 5 payments
```{r}
dbGetQuery(con, "
SELECT 
  customer_id,
  MAX(amount) AS `max`,
  MIN(amount) AS `min`,
  AVG(amount) AS `avg`,
  SUM(amount) AS `sum`
FROM payment
GROUP BY customer_id
HAVING COUNT(*) >5")
```

# Clean up
```{r}
dbDisconnect(con)
```


