---
title: "LAB3"
author: "Caroline"
date: "9/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library data.table
#Install if and only if it's not present
```{r}
if (!require(data.table)){
  install.packages("data.table")
}
```

## Step 1 Download and read the dataset
#Install if we don't have it.
```{r}
if (!file.exists("met_all.gz")) {
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
              destfile = "met_all.gz", 
              method="libcurl", 
              timeout = 60
              )
}

met <- data.table::fread("met_all.gz")
```

## Step 2 Check the dimensions, headers, footers, rows and columns 
Dimensions
```{r}
dim(met)
```
Headers
```{r}
head(met)
```
Footers
```{r}
tail(met)
```
Rows
```{r}
nrow(met)
```
Columns
```{r}
ncol(met)
```

## Step 3 Brief look at the variables
```{r}
str(met)
```

## Step 4 Closer look at the key variables

```{r}
table(met$year)

table(met$day)

table(met$hour)

summary(met$temp)

summary(met$elev)
```

update NAs in elevation
```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```

remove temperature of -40C
```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```

## Step 5 Check the data against an external data source
```{r}
#make the temperature reasonable (temp > -15C)
met <- met[temp>-15][order(temp)]
summary(met[, .(lat, lon, wind.sp, temp, elev)])
head(met)
```

## Step 6 Calculate summary statistics
```{r}
#select station with minimum elevation
elev <- met[elev==min(elev)]
summary(elev)
```
correlation between temperature and wind speed
```{r}
cor(elev$temp, elev$wind.sp, use="complete")
```
correlation between temperature and hours
```{r}
cor(elev$temp, elev$hour, use="complete")
```
correlation between wind speed and day
```{r}
cor(elev$wind.sp, elev$day, use="complete")
```
correlation between wind speed and hour
```{r}
cor(elev$wind.sp, elev$hour, use="complete")
```
correlation between temperature and day
```{r}
cor(elev$temp, elev$day, use="complete")
```

## Step 7 Exploratory Graphs
