---
title: "LAB4"
author: "Caroline"
date: "9/17/2021"
output: github_document
always_allow_html: TRUE
---

## R package setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(data.table)
library(tidyverse)
```


## Step 1: Read in the data
```{r}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
```

## Step 2: Prepare the data
```{r}
#Remove temperatures less than -17C
met <- met[temp >= -17]

#Make sure there are no missing data in the key variables coded as 9999, 999, etc
#temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev
met[,range(temp)]
met[,range(rh, na.rm = TRUE)]
met[,range(wind.sp, na.rm = TRUE)]
met[,range(vis.dist, na.rm = TRUE)]
met[,range(dew.point, na.rm = TRUE)]
met[,range(lat, na.rm = TRUE)]
met[,range(lon, na.rm = TRUE)]
met[,range(elev, na.rm = TRUE)]

met[elev == 9999.0, elev := NA]

#Generate a date variable using the functions as.Date() (hint: You will need the following to create a date paste(year, month, day, sep = "-")).
met[, ymd := as.Date(paste(year, month, day, sep = "-"))]

#Using the data.table::week function, keep the observations of the first week of the month.
met[, table(week(ymd))]
met <- met[ week(ymd) == 31]

#Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.
met_avg <- met[, .(
  temp = mean(temp, na.rm = TRUE),
  rh = mean(rh, na.rm = TRUE),
  wind.sp = mean(wind.sp, na.rm = TRUE),
  vis.dist = mean(vis.dist, na.rm = TRUE),
  dew.point = mean(dew.point, na.rm = TRUE),
  lat = mean(lat, na.rm = TRUE),
  lon = mean(lon, na.rm = TRUE),
  elev = mean(elev, na.rm = TRUE), USAFID
), by = "USAFID"]

#Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
met_avg[lat >= 39.71 & lon <= -98, region := "Northwest"]
met_avg[lat < 39.71 & lon <= -98, region := "Southwest"]
met_avg[lat >= 39.71 & lon > -98, region := "Northeast"]
met_avg[lat < 39.71 & lon > -98, region := "Southeast"]

met_avg[, region2 := fifelse(lat >= 39.71 & lon <= 98, "Northwest",
        fifelse(lat < 39.71 & lon <= -98, "Southwest",
          fifelse(lat >= 39.71 & lon > -98, "Northeast",
            fifelse(lat < 39.71 & lon > -98, "Southeast", NA_character_))))]

met_avg[, table(region, useNA = "always")]

#Create a categorical variable for elevation as in the lecture slides
met_avg[, elev_cat := fifelse(elev > 252, "high", "low")]

#Delete extra column
met_avg <- met_avg[, -1]
```

## Step 3: Use geom_violin to examine the wind speed and dew point temperature by region
```{r violin-temp}
ggplot(met_avg, mapping = aes(y = wind.sp, x = 1)) +
  geom_violin() + 
  facet_grid(~region)
```

## Step 4: Use geom_jitter with stat_smooth to examine the association between dew point temperature and wind speed by region
```{r}
met_avg[!is.na(dew.point) & !is.na(wind.sp)] %>%
  ggplot(mapping = aes(x = dew.point, y = wind.sp, color = region, linetype = region)) +
  geom_jitter() +
  stat_smooth(method = "lm", se = FALSE, color = "blue")
```
The wind speed remained almost the same as the dew. point changed.


## Step 5: Use geom_bar to create barplots of the weather stations by elevation category coloured by region
```{r}
met_avg[!is.na(elev_cat)] %>%
    ggplot() + 
    geom_bar(mapping = aes(x = elev_cat, fill = region), position="dodge") + 
    labs(
      x = "elevation category",
      y = "frequency",
      title = "barplot of the weather station by elevation category") +
    scale_fill_brewer(palette = "GnBu")
```
The southeast with low elevation show the highest frequency

## Step 6: Use stat_summary to examine mean dew point and wind speed by region with standard deviation error bars

```{r ploy-stat sum}
ggplot(
  met_avg[!is.na(wind.sp) & !is.na(dew.point)],
  mapping = aes(x = region, y = wind.sp)) +
  stat_summary(fun.data = "mean_sdl") +
  stat_summary(fun.data = "mean_sdl", geom = "errorbar") 
```

## Step 7: Make a map showing the spatial trend in relative h in the US
```{r leaflet}
library(leaflet)

temp.pal <- colorNumeric(c('light sky blue','goldenrod','brown'),
                         domain = met_avg$rh)

leaflet(met_avg) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addCircles(
    lat = ~lat, lng=~lon,
     label = ~paste0(round(temp,2), ' C'), color = ~ temp.pal(temp),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  addLegend('bottomleft', pal=temp.pal, values=met_avg$temp,
          title='Temperature, C', opacity=1)
```

## Step 8: Use a ggplot extension
```{r}
library(gghalves)
ggplot(met_avg, aes(wind.sp, temp)) +
  geom_boxplot() +
  geom_half_violin(side = "l")
```

