---
title: "LAB7"
author: "Caroline He"
date: "10/8/2021"
output: 
  github_document: default
  html_document:
    html_preview: false
  word_document: default 
always_allow_html: true
---

# Web scraping and Regular Expressions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Learning Goals
* Use a real world API to make queries and process the data.

* Use regular expressions to parse the information.

* Practice your GitHub skills.

## Lab description
In this lab, we will be working with the NCBI API to make queries and extract information using XML and regular expressions. For this lab, we will be using the httr, xml2, and stringr R packages.

```{r}
#r packages
library(tidyverse)
library(tidytext)
library(stringr)
library(httr)
```


### Question 1: How many sars-cov-2 papers?
Build an automatic counter of sars-cov-2 papers using PubMed. You will need to apply XPath as we did during the lecture to extract the number of results returned by PubMed in the following web address:https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2

```{r}
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2")

# Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/section[1]/div[2]/div[1]/span")

# Turning it into text
counts <- as.character(counts)

# Extracting the data using regex
stringr::str_extract(counts, "[0-9,]+")
stringr::str_extract(counts, "[[:digit:],]+")
```

### Question 2: Academic publications on COVID19 and Hawaii
You need to query the following The parameters passed to the query are documented here.

Use the function httr::GET() to make the following query:

1. Baseline URL: https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi
2. Query parameters:

* db: pubmed
* term: covid19 hawaii
* retmax: 1000

```{r}
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db     = "pubmed",
    term   = "covid19 hawaii",
    retmax = 1000
    )
)

# Extracting the content of the response of GET
ids <- httr::content(query_ids)
ids
```

The query will return an XML object, we can turn it into a character list to analyze the text directly with as.character(). Another way of processing the data could be using lists with the function xml2::as_list(). We will skip the latter for now.

### Question 3: Get details about the articles
The Ids are wrapped around text in the following way: <Id>... id number ...</Id>. we can use a regular expression that extract that information. 
```{r}
# Turn the result into a character vector
ids <- as.character(ids)

# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[[:digit:]]+</Id>")[[1]]

# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- stringr::str_remove_all(ids, "<Id>|</Id>")
ids
```

With the ids in hand, we can now try to get the abstracts of the papers. As before, we will need to coerce the contents (results) to a list using:

1. Baseline url: https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi
2. uery parameters:

* db: pubmed
* id: A character with all the ids separated by comma, e.g., “1232131,546464,13131”
* retmax: 1000
* rettype: abstract

Pro-tip: If you want GET() to take some element literal, wrap it around I() (as you would do in a formula in R). For example, the text "123,456" is replaced with "123%2C456". If you don’t want that behavior, you would need to do the following I("123,456").

```{r}
publications <- GET(
  url = "https://eutils.ncbi.nlm.nih.gov/",
  path = "entrez/eutils/efetch.fcgi",
  query = list(
    db = "pubmed",
    id = paste(ids, collapse = ","),
    retmax = 1000,
    rettype = "abstract"
    )
)

# Turning the output into character vector
publications <- httr::content(publications)
publications_txt <- as.character(publications)
```

### Question 4: Distribution of universities, schools, and departments
Using the function stringr::str_extract_all() applied on publications_txt, capture all the terms of the form:

1. University of …
2. … Institute of …

Write a regular expression that captures all such instances

```{r}
institution <- str_extract_all(
  str_to_lower(publications_txt),
  "university\\s+of\\s+(southern|new|northern|the)?\\s+w[[:alpha:]-]+|[[:alpha:]-]+\\s+institute\\s+of\\s+[[:alpha:]-]+"
  ) 
institution <- unlist(institution)
table(institution)
```

Repeat the exercise and this time focus on schools and departments in the form of

1. School of …
2. Department of …

And tabulate the results
```{r}
schools_and_deps <- str_extract_all(
  str_to_lower(publications_txt),
  "school\\s+of\\s+(southern|new|northern|the)?\\s+w[[:alpha:]-]+|department\\s+of\\s+[[:alpha:]-]+"
  )

table(schools_and_deps)
```

### Question 5: Form a database
We want to build a dataset which includes the title and the abstract of the paper. The title of all records is enclosed by the HTML tag ArticleTitle, and the abstract by Abstract.

Before applying the functions to extract text directly, it will help to process the XML a bit. We will use the xml2::xml_children() function to keep one element per id. This way, if a paper is missing the abstract, or something else, we will be able to properly match PUBMED IDS with their corresponding records.
```{r}
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
```

extract the abstract and article title for each one of the elements of pub_char_list. You can either use sapply() as we just did, or simply take advantage of vectorization of stringr::str_extract
```{r}
abstracts <- str_extract(pub_char_list, "<Abstract>[[:print:][:space:]]+</Abstract>")
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]- =\"]+>")
abstracts <- str_replace_all(abstracts, "[[:space:]]+", " ")
```

How many of these don’t have an abstract? Now, the title
```{r}
titles <- str_extract(pub_char_list, "<ArticleTitle>[[:print:][:space:]]+</ArticleTitle>")
titles <- str_remove_all(titles, "</?[[:alnum:]- =\"]+>")
```

Finally, put everything together into a single data.frame and use knitr::kable to print the results
```{r}
database <- data.frame(
  PubMedId = ids,
  Title    = titles,
  Abstract = abstracts
)
knitr::kable(database[1:20,], caption = "Some papers about Covid19 and Hawaii")
```



